Load original data.

```{r, warning=FALSE, echo=FALSE}

library(readxl)
library(tidyverse)

cpr <- readxl::read_excel(path="data/chestPainRegistry.xlsx",
                          na=c("9", "999", "9999", "6666")) |> 
  dplyr::rename_all(tolower)
```

Select patients with acs, and variables of interest. Create a variable to split cohort into train and test subgroups.

```{r, warning=FALSE}

acs <- cpr |> 
  filter(!(is.na(`ficha resca`))) |> 
  mutate_if(is.character, as.double) |> 
  transmute(mace = ifelse(obitot == 1 | iamcsst == 1 | iamssst == 1, 1, 0),
            majorBleeding = sangmaior,
            age = idade,
            maleSex = sexo,
            bmi = peso / alt^2,
            currentSmoker = taba,
            dyslipidemia = disl,
            diabetesMellitus = dm,
            hypertension = has,
            kidneyDisease = irc,
            peripheralDisease = dapp,
            carotidDisease = carotidap,
            coronaryDisease = dacp...27,
            familyHistory = hfdac,
            priorMi = iamp,
            priorStroke = avcp,
            priorAspirin = aasp,
            pulse = fc,
            systolicBp = pas,
            killip = ifelse(killip != 1, 1, 0),
            ecgChanges = ifelse(ecgisq == 1, 1, 0),
            myocardialInjury = tnalta,
            ntProBnp = probnp,
            creatinine = creata,
            hemoglobin = hb1)

# Label variables
varLabels <- c("Eventos cardiovasculares adversos graves",
               "Sangramento grave",
               "Idade (anos)",
               "Sexo masculino",
               "Indice de massa corporal",
               "Tabagismo atual",
               "Dislipidemia",
               "Diabetes mellitus",
               "Hipertensão arterial sistêmica",
               "Doença renal crônica",
               "Doença arterial periférica",
               "Doença arterial carotídea",
               "Doença arterial coronariana",
               "História familiar de DAC prematura",
               "Infarto prévio",
               "Acidente vascular encefálico prévio",
               "Uso prévio de AAS",
               "Frequência cardíaca (min⁻¹)",
               "Pressão arterial sistólica (mmHg)",
               "Classe de Killip II-IV",
               "Alterações isquêmicas no ECG",
               "Injúria miocárdica",
               "NT pró-BNP (pg.mL⁻¹)",
               "Creatinina (mg.dL⁻¹)",
               "Hemoglobina (g.dL⁻¹)")

library(labelled)
var_label(acs) <- varLabels

# Create variable to split into train and test subgroups
set.seed(1)
acs$trainSet <- ifelse(runif(nrow(acs)) < 0.8, TRUE, FALSE)

# Save and see results
save(acs, file = "data/acs.rda")

head(acs)
str(acs)
```

Describe baseline characteristics, comparing train and test subgroups (table).

```{r, warning=FALSE}

library(gt)
library(gtsummary)

theme_gtsummary_language(language = "pt",
                         decimal.mark = ",",
                         big.mark = ".",
                         iqr.sep = "–")

baselineChars <- acs |> 
  select(-c(mace, majorBleeding)) |> 
  gtsummary::tbl_summary(by = trainSet,
                         missing="no",
                         statistic=list(all_categorical()~"{n}/{N} ({p}%)",
                                        all_continuous()~"{mean} ± {sd}",
                                        c("ntProBnp", "creatinine")~
                                          "{median} ({p25}-{p75})")) |> 
  add_stat_label(label=list(all_categorical()~"n/N (%)",
                            all_continuous()~"média (DP)",
                            c("ntProBnp", "creatinine")~"mediana (IIQ)")) |> 
  modify_header(p.value="**P**",
                label="**Característica**",
                update = list(stat_1 ~ "**Validação** (N = {n})",
                              stat_2 ~ "**Derivação** (N = {n})")) |> 
  modify_spanning_header(c("stat_1", "stat_2") ~ c("**Subgrupos**")) |> 
  add_overall() |>
  modify_footnote(update=everything()~NA) |>
  as_gt() |> 
  opt_table_font(font = google_font("Times"))

baselineChars
gtsave(baselineChars, filename = "tables/baselineChars.png")

```

Split cohort into train and test subsets. Impute each subset.

```{r, warning=FALSE}

# Split cohort
train <- filter(acs, trainSet == TRUE)
test <- filter(acs, trainSet == FALSE)

# Remove "subgroup" variable from subsets
train$trainSet <- NULL
test$trainSet <- NULL

# Create predictor matrix
p <- ncol(train)
pMatrix <- matrix(rep(1, p^2),
                  nrow = p,
                  ncol = p)
diag(pMatrix) <- 0

# Define imputation method for each predictor
dput(names(train))
impMethod <- c("logreg", "logreg", "pmm", "logreg", "pmm", "logreg", 
               "logreg", "logreg", "logreg", "logreg", "logreg", "logreg",
               "logreg", "logreg", "logreg", "logreg", "logreg", "pmm", 
               "pmm", "logreg", "logreg", "logreg", "pmm", "pmm", "pmm")

# Impute each subset
library(mice)

impTrain <- mice(train,
                 m = 40,
                 maxit = 100,
                 predictorMatrix = pMatrix,
                 method = impMethod)
impTest <- mice(test,
                m = 40,
                maxit = 100,
                predictor_matrix = pMatrix,
                method = impMethod,
                seed = 1)

save(impTrain, file = "imputation/train.rda")
save(impTest, file = "imputation/test.rda")

# See results

load("imputation/train.rda")
load("imputation/test.rda")

impTrain
impTest
```

Diagnostics for missing data and multiple imputation (fig).

```{r, warning=FALSE}

# Packages
library(Hmisc)
library(VIM)
library(mice)

# Diagnostics for missing data
dput(names(acs))
predVars <- c("age", "maleSex", "bmi", "currentSmoker",
              "dyslipidemia", "diabetesMellitus", "hypertension",
              "kidneyDisease", "peripheralDisease", "carotidDisease",
              "coronaryDisease", "familyHistory", "priorMi", "priorStroke",
              "priorAspirin", "pulse", "systolicBp", "killip", "ecgChanges",
              "myocardialInjury", "ntProBnp", "creatinine", "hemoglobin")
predLabels <-
  acs[predVars] |> 
  labelled::var_label(unlist = TRUE) |> 
  as.vector() |> 
  stringr::str_replace("\\([:graph:]*", "") |>
  stringr::str_trim()

# Missing values per variable combination
acs [predVars] |> 
  VIM::aggr(labels = predLabels,
            col = c("#6657fe", "#fd0a45"),
            combined = TRUE,
            sortVars = TRUE,
            sortCombs = TRUE,
            oma = c(24, 1, 1, 1),
            cex.axis = 1,
            ylabs = "")

# Hierarchical cluster analysis of missing values
acs[predVars] |> 
  Hmisc::naclus() |>
  plot(ylab = "Fração dos Valores Faltantes em Comum")

# Number of missing values per variable
acs[predVars] |> 
  Hmisc::naclus() |>
  naplot(which="na per var")

# Number of missing values per observation
acs[predVars] |> 
  Hmisc::naclus() |>
  naplot(which="na per obs")

# Number of other variables missing, when the indicated variable is missing
acs[predVars] |> 
  Hmisc::naclus() |>
  naplot(which="mean na")

# Trace lines of mice algorithm
load("imputation/train.rda")
plot(impTrain, 
     layout = c(4, 12)) |> 
  update(main = "Linhas de rastreio do algoritmo de imputação",
         xlab = "Índice de iteração",
         ylab = "Valor estimado")

load("imputation/test.rda")
plot(impTest, 
     layout = c(4, 12)) |> 
  update(main = "Linhas de rastreio do algoritmo de imputação",
         xlab = "Índice de iteração",
         ylab = "Valor estimado")

# Compare observed and imputed data
densityplot(impTrain,
            ~bmi+pulse+systolicBp+ntProBnp+creatinine+hemoglobin) |> 
  update(main = "Estimativas de Densidade Kernel",
         xlab = "Valores da variável",
         ylab = "Densidade")
densityplot(impTest,
            ~bmi+pulse+systolicBp+ntProBnp+creatinine+hemoglobin) |> 
  update(main = "Estimativas de Densidade Kernel",
         xlab = "Valores da variável",
         ylab = "Densidade")
```

Regression: sample size calculation (test).

```{r, warning=FALSE}

# Rule of thumb: 10 events per predictor parameter
nPredVars <- function(nEvents, epp){
  nEvents / epp
}

# Train subset
train <- filter(acs, trainSet == TRUE)

# Number of candidate predictors for mace
nPredVars(nEvents = sum(train$mace, na.rm = TRUE),
          epp = 10)

# Number of candicate predictors for major bleeding
nPredVars(nEvents = sum(train$majorBleeding, na.rm = TRUE),
          epp = 10)
```

Regression: perform univariate analysis to select predictors (table).

```{r, warning=FALSE}

library(gt)
library(gtsummary)

theme_gtsummary_language(language = "pt",
                         decimal.mark = ",",
                         big.mark = ".",
                         iqr.sep = "–")

# Associations with mace
acs |> 
  filter(trainSet == TRUE) |> 
  mutate(trainSet = NULL) |>
  select(-c(majorBleeding)) |> 
  tbl_summary(by = mace,
              missing = "no",
              statistic = list(all_categorical()~"{n}/{N} ({p}%)",
                                        all_continuous()~"{mean} ± {sd}",
                                        c("ntProBnp", "creatinine")~
                                          "{median} ({p25}-{p75})")) |>
  add_stat_label(label=list(all_categorical()~"n/N (%)",
                            all_continuous()~"média (DP)",
                            c("ntProBnp", "creatinine")~"mediana (IIQ)")) |>
  modify_header(p.value="**P**",
                label="**Característica**",
                update = list(stat_1 ~ "**Não** (N = {n})",
                              stat_2 ~ "**Sim** (N = {n})")) |> 
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 3)) |> 
  modify_spanning_header(c("stat_1", "stat_2")~"**Óbito ou Reinfarto**") |> 
  modify_footnote(update = everything() ~ NA) |> 
  bold_p(t = 0.005) |> 
  as_gt() |> 
  opt_table_font(font = google_font("Times")) |> 
  gtsave(filename = "tables/predictorsMace.png")

# Associations with major bleeding
acs |> 
  filter(trainSet == TRUE) |> 
  mutate(trainSet = NULL) |>
  select(-c(mace)) |> 
  tbl_summary(by = majorBleeding,
              missing = "no",
              statistic = list(all_categorical()~"{n}/{N} ({p}%)",
                                        all_continuous()~"{mean} ± {sd}",
                                        c("ntProBnp", "creatinine")~
                                          "{median} ({p25}-{p75})")) |>
  add_stat_label(label=list(all_categorical()~"n/N (%)",
                            all_continuous()~"média (DP)",
                            c("ntProBnp", "creatinine")~"mediana (IIQ)")) |>
  modify_header(p.value="**P**",
                label="**Característica**",
                update = list(stat_1 ~ "**Não** (N = {n})",
                              stat_2 ~ "**Sim** (N = {n})")) |> 
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 3)) |> 
  modify_spanning_header(c("stat_1", "stat_2")~"**Sangramento Maior**") |> 
  modify_footnote(update = everything() ~ NA) |> 
  bold_p(t = 0.001) |> 
  as_gt() |> 
  opt_table_font(font = google_font("Times")) |> 
  gtsave(filename = "tables/predictorsBleeding.png")
```

Regression: derive model and validate predictions.

```{r, warning=FALSE}

# Load rms package to perform logistic regression
library(rms)

# Load imputed train and test subgroups
load("imputation/train.rda")
load("imputation/test.rda")

# Load original train and test subgroups
train <- filter(acs, trainSet == TRUE)
test <- filter(acs, trainSet == FALSE)

# Look at variable names to select predictors
dput(names(acs)) |> paste(collapse = "+")

# Specify datadist for logistic regression
dd <- datadist(train)
options(datadist = "dd")

# Derive logistic regression for mace
lrmMace <- fit.mult.impute(mace~age+kidneyDisease+pulse+killip+ntProBnp+creatinine+hemoglobin,
                           fitter = glm,
                           xtrans = impTrain,
                           data = train) |> 
  step()
lrmMace <- fit.mult.impute(lrmMace$formula,
                           fitter = lrm,
                           xtrans = impTrain,
                           data = train)
save(lrmMace, file="models/lrmMace.rda")

# Validate logistic regression for mace
predictionLrmMace <- predict(lrmMace,
                             newdata = complete(impTest, 1),
                             type = "fitted")
save(predictionLrmMace, file="models/predictionLrmMace.rda")


# Derive logistic regression for major bleeding
lrmBleeding <- fit.mult.impute(majorBleeding~age+killip+myocardialInjury+ntProBnp+hemoglobin,
                           fitter = glm,
                           xtrans = impTrain,
                           data = train) |> 
  step()
lrmBleeding <- fit.mult.impute(lrmBleeding$formula,
                           fitter = lrm,
                           xtrans = impTrain,
                           data = train)
save(lrmBleeding, file="models/lrmBleeding.rda")

# Validate logistic regression for major bleeding
predictionLrmBleeding <- predict(lrmBleeding,
                             newdata = complete(impTest, 1),
                             type = "fitted")
save(predictionLrmBleeding, file="models/predictionLrmBleeding.rda")
```

SVM: train model, assess variable importance, and validate predictions.

```{r, warning=FALSE}


# Load caret to perform svm
library(caret)

# Set parameters for model training
trainControlSvm <- trainControl(method = "repeatedcv",
                                number = 10,
                                repeats = 100,
                                classProbs = TRUE)

# Train svm for predicting mace
svmMace <- train(make.names(mace) ~ .,
                 data = select(complete(impTrain, 1),
                               -c(majorBleeding)),
                 method = "svmRadial",
                 trControl = trainControlSvm,
                 metric =  "ROC",
                 tuneLength = 10,
                 preProcess = c("center", "scale"))
save(svmMace, file="models/svmMace.rda")

# Variable importance of svm for predicting mace
varImpSvmMace <- varImp(svmMace, scale = FALSE)
save(varImpSvmMace, file="models/varImpSvmMace.rda")

# Validate mace predictions
load("models/svmMace.rda")
load("imputation/test.rda")

predictionSvmMace <- predict(svmMace, 
                             newdata = complete(impTest, 1),
                             type="prob")
save(predictionSvmMace, file="models/predictionSvmMace.rda")

# Train svm for predicting major bleeding
svmBleeding <- train(make.names(majorBleeding) ~ .,
                 data = select(complete(impTrain, 1),
                               -c(mace)),
                 method = "svmRadial",
                 trControl = trainControlSvm,
                 metric =  "ROC",
                 tuneLength = 10,
                 preProcess = c("center", "scale"))
save(svmBleeding, file="models/svmBleeding.rda")

# Variable importance of svm for predicting major bleeding
varImpSvmBleeding <- varImp(svmBleeding, scale = FALSE)
save(varImpSvmBleeding, file="models/varImpSvmBleeding.rda")

# Validate bleeding predictions
predictionSvmBleeding <- predict(svmBleeding, 
                             newdata = complete(impTest, 1),
                             type="prob")
save(predictionSvmBleeding, file="models/predictionSvmBleeding.rda")
```

XGBoost: train model, assess variable importance, and validate predictions.

```{r, warning=FALSE}

# Load caret to perform xgb
library(caret)

# Set parameters for model training
trainControlXgb <- trainControl(method = "cv",
                                number = 10,
                                classProbs = TRUE,
                                allowParallel = TRUE)

# Train xgb for predicting mace
xgbMace <- train(make.names(mace) ~ .,
                 data = select(complete(impTrain, 1),
                               -c(majorBleeding)),
                 method = "xgbTree",
                 preProcess = c("center", "scale"),
                 trControl = trainControlXgb,
                 tuneLength = 10,
                 verbose = TRUE,
                 metric =  "ROC")
save(xgbMace, file="models/xgbMace.rda")

# Variable importance of svm for predicting mace
varImpXgbMace <- varImp(xgbMace, scale = FALSE)
save(varImpXgbMace, file="models/varImpXgbMace.rda")

# Validate mace predictions
predictionXgbMace <- predict(xgbMace, 
                             newdata = complete(impTest, 1),
                             type="prob")
save(predictionXgbMace, file="models/predictionXgbMace.rda")

# Train svm for predicting major bleeding
xgbBleeding <- train(make.names(majorBleeding) ~ .,
                 data = select(complete(impTrain, 1),
                               -c(mace)),
                 method = "xgbTree",
                 preProcess = c("center", "scale"),
                 trControl = trainControlXgb,
                 tuneLength = 10,
                 verbose = TRUE,
                 metric =  "ROC")
save(xgbBleeding, file="models/xgbBleeding.rda")

# Variable importance of svm for predicting major bleeding
varImpXgbBleeding <- varImp(xgbBleeding, scale = FALSE)
save(varImpXgbBleeding, file="models/varImpXgbBleeding.rda")

# Validate bleeding predictions
predictionXgbBleeding <- predict(xgbBleeding, 
                             newdata = complete(impTest, 1),
                             type="prob")
save(predictionXgbBleeding, file="models/predictionXgbBleeding.rda")
```

Logistic regression results for predicting MACE (fig).

```{r, warning=FALSE}

# Packages
library(tidyverse)
library(gridExtra)
library(rms)

# Dataframe containg results
load("models/lrmMace.rda")

dfOrMace <- data.frame(label = c("Intercept",
                                 lrmMace[["Design"]][["label"]]),
                       beta = lrmMace$coefficients,
                       se = sqrt(diag(lrmMace$var)),
                       row.names = NULL) |>
  filter(label != "Intercept") |> 
  mutate(or = exp(beta),
         ll = or + qnorm(0.025)*se,
         ul = or + qnorm(0.975)*se,
         z = beta/se,
         p = 1 - pchisq(z^2, 1),
         p = ifelse(p < 0.001, "< 0,001", as.character(round(p, 3))))

# Forest plot
forestMace <- ggplot(dfOrMace,
       aes(x = or, 
           y = reorder(label, or))) +
  geom_vline(xintercept = 1,
             color = "black",
             linetype = "dashed",
             cex = 0.5,
             alpha = 0.5) +
  geom_errorbarh(aes(xmin = ll,
                     xmax = ul),
                 height = 0) +
  geom_point(shape = 15, 
             size = 5,
             color = "#850A70") +
  scale_x_continuous(limits = c(0, 4),
                breaks = c(0, 2, 4),
                expand = expansion(0, 0)) +
  labs(y = "",
       subtitle = "Razão de Chances (IC 95%)\n") +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.length.x = unit(7.5, "pt"),
        axis.line.x = element_line(color = "black"),
        axis.text.y = element_text(size = 12,
                                   color = "black",
                                   family = "Montserrat",
                                   margin = margin(r = 20),
                                   hjust = 0),
        axis.text.x.bottom = element_text(size = 12,
                                          color = "black",
                                          family = "Montserrat",
                                          margin = margin(t = 10, b = 20)),
        axis.title.x = element_blank(),
        plot.subtitle = element_text(size = 12,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

# Base table
tableBaseMace <- ggplot(dfOrMace, aes(y=label)) +
  labs(x = "",
       y = NULL) +
  theme(plot.title = element_text(hjust = 0.5,
                                  size=12), 
        axis.text.x = element_text(color="white", 
                                   hjust = -3, 
                                   size = 25),
        axis.line = element_blank(),
        axis.text.y = element_blank(), 
        axis.ticks = element_blank(),
        axis.title.y = element_blank(), 
        legend.position = "none",
        panel.background = element_blank(), 
        panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        plot.background = element_blank())

# Point estimate table
tableEstimateMace <- tableBaseMace + 
  labs(title = "\n") +
  geom_text(aes(y = reorder(label, or),
                x = 1,
                label = paste(sprintf("%0.2f",
                                      round(or, 1)),
                              " (",
                              sprintf("%0.2f",
                                      round(ll, 1)),
                              "-",
                              sprintf("%0.2f",
                                      round(ul, 1)),
                              ")", 
                              sep = "")),
            size = 4,
            family = "Montserrat",
            hjust = 0.5) + ## decimal places
  theme(title = element_text(size = 12,
                             color = "black",
                             family = "Montserrat",
                             face = "bold"))

# P value table
tablePMace <- tableBaseMace + 
  labs(title = "P\n") +
  geom_text(aes(y = reorder(label, or),
                x = 1,
                label = p),
            size = 4,
            color = "black",
            family = "Montserrat",
            hjust = 0.8) +
  theme(title = element_text(size = 12,
                             color = "black",
                             family = "Montserrat",
                             face = "bold"))


# combining elements of forest plot
tableLayout <-  matrix(c(1,1,1,1,1,1,2,2,2,3,3), nrow = 1)
grid.arrange(forestMace, tableEstimateMace, tablePMace, layout_matrix = tableLayout)
```

Logistic regression results for predicting major bleeding (fig).

```{r, warning=FALSE}

# Packages
library(tidyverse)
library(gridExtra)
library(rms)

# Dataframe containg results
load("models/lrmBleeding.rda")

dfOrBleeding <- data.frame(label = c("Intercept",
                                 lrmBleeding[["Design"]][["label"]]),
                       beta = lrmBleeding$coefficients,
                       se = sqrt(diag(lrmBleeding$var)),
                       row.names = NULL) |>
  filter(label != "Intercept") |> 
  mutate(or = exp(beta),
         ll = or + qnorm(0.025)*se,
         ul = or + qnorm(0.975)*se,
         z = beta/se,
         p = 1 - pchisq(z^2, 1),
         p = ifelse(p < 0.001, "< 0,001", as.character(round(p, 3))))

# Forest plot
forestBleeding <- ggplot(dfOrBleeding,
       aes(x = or, 
           y = reorder(label, or))) +
  geom_vline(xintercept = 1,
             color = "black",
             linetype = "dashed",
             cex = 0.5,
             alpha = 0.5) +
  geom_errorbarh(aes(xmin = ll,
                     xmax = ul),
                 height = 0) +
  geom_point(shape = 15, 
             size = 5,
             color = "#257F93") +
  scale_x_continuous(limits = c(0, 5),
                breaks = c(0, 2.5, 5),
                expand = expansion(0, 0)) +
  labs(y = "",
       subtitle = "Razão de Chances (IC 95%)\n") +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.length.x = unit(7.5, "pt"),
        axis.line.x = element_line(color = "black"),
        axis.text.y = element_text(size = 12,
                                   color = "black",
                                   family = "Montserrat",
                                   margin = margin(r = 20),
                                   hjust = 0),
        axis.text.x.bottom = element_text(size = 12,
                                          color = "black",
                                          family = "Montserrat",
                                          margin = margin(t = 10, b = 20)),
        axis.title.x = element_blank(),
        plot.subtitle = element_text(size = 12,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

# Base table
tableBaseBleeding <- ggplot(dfOrBleeding, aes(y=label)) +
  labs(x = "",
       y = NULL) +
  theme(plot.title = element_text(hjust = 0.5,
                                  size=12), 
        axis.text.x = element_text(color="white", 
                                   hjust = -3, 
                                   size = 25),
        axis.line = element_blank(),
        axis.text.y = element_blank(), 
        axis.ticks = element_blank(),
        axis.title.y = element_blank(), 
        legend.position = "none",
        panel.background = element_blank(), 
        panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        plot.background = element_blank())

# Point estimate table
tableEstimateBleeding <- tableBaseBleeding + 
  labs(title = "\n") +
  geom_text(aes(y = reorder(label, or),
                x = 1,
                label = paste(sprintf("%0.2f",
                                      round(or, 1)),
                              " (",
                              sprintf("%0.2f",
                                      round(ll, 1)),
                              "-",
                              sprintf("%0.2f",
                                      round(ul, 1)),
                              ")", 
                              sep = "")),
            size = 4,
            family = "Montserrat",
            hjust = 0.5) + 
  theme(title = element_text(size = 12,
                             color = "black",
                             family = "Montserrat",
                             face = "bold"))

# P value table
tablePBleeding <- tableBaseBleeding + 
  labs(title = "P\n") +
  geom_text(aes(y = reorder(label, or),
                x = 1,
                label = p),
            size = 4,
            color = "black",
            family = "Montserrat",
            hjust = 0.8) +
  theme(title = element_text(size = 12,
                             color = "black",
                             family = "Montserrat",
                             face = "bold"))

# combining elements of forest plot
tableLayout <-  matrix(c(1,1,1,1,1,1,2,2,2,3,3), nrow = 1)
grid.arrange(forestBleeding, tableEstimateBleeding, tablePBleeding, layout_matrix = tableLayout)
```

Variable importance in machine learning predictions (fig).

```{r, warning=FALSE}

# Variable labels
load("data/acs.rda")

predVars <- c("age", "maleSex", "bmi", "currentSmoker",
              "dyslipidemia", "diabetesMellitus", "hypertension",
              "kidneyDisease", "peripheralDisease", "carotidDisease",
              "coronaryDisease", "familyHistory", "priorMi", "priorStroke",
              "priorAspirin", "pulse", "systolicBp", "killip", "ecgChanges",
              "myocardialInjury", "ntProBnp", "creatinine", "hemoglobin")
predLabels <-
  acs[predVars] |> 
  labelled::var_label(unlist = TRUE) |> 
  as.vector() |> 
  stringr::str_replace("\\([:graph:]*", "") |>
  stringr::str_trim()

# Dataframe of variable importance estimates
load("models/varImpSvmMace.rda")
load("models/varImpXgbMace.rda")
load("models/varImpSvmBleeding.rda")
load("models/varImpXgbBleeding.rda")

dfVarImp <- data.frame(label = predLabels |> 
                         str_replace("\\([:graph:]*", ""),
                       svmMace = varImpSvmMace$importance$Overall,
                       svmBleeding = varImpSvmBleeding$importance$Overall,
                       xgbMace = varImpXgbMace$importance$Overall,
                       xgbBleeding=varImpXgbBleeding$importance$Overall) |>
  mutate(scaledSvmMace = svmMace / max(svmMace) * 100,
         scaledSvmBleeding = svmBleeding / max(svmBleeding) * 100,
         scaledXgbMace = xgbMace / max(xgbMace) * 100,
         scaledXgbBleeding = xgbBleeding / max(xgbBleeding) * 100,
         meanMace = (scaledSvmMace + scaledXgbMace) / 2,
         meanBleeding = (scaledSvmBleeding + scaledXgbBleeding) / 2)

# Importance for MACE
varImpMace <- ggplot(dfVarImp, 
       aes(y = reorder(label, meanMace))) +
  geom_segment(aes(y = reorder(label, meanMace), 
                   yend = reorder(label, meanMace),
                   x = scaledSvmMace,
                   xend = scaledXgbMace),
               color = "lightgrey") +
  geom_point(aes(x = scaledSvmMace,
                 color = "SVM"),
             alpha = 0.7) +
  geom_point(aes(x = scaledXgbMace,
                 color = "XGB"),
             alpha = 0.7) +
  scale_y_discrete(position = "right") +
  scale_x_continuous(position = "top") +
  labs(x = "Estimativa Dimensionada de Desempenho (%)\n") +
  scale_color_manual(name = "Modelo preditor de\neventos cardiovasculares",
                     breaks = c("SVM", "XGB"),
                     values = c("SVM" = "#009fc3", "XGB" = "#6cb43f")) +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 11,
                                   color = "black",
                                   family = "Montserrat",
                                   face = "bold"),
        axis.text.y = element_text(size = 10,
                                   color = "black",
                                   family = "Montserrat",
                                   margin = margin(r = 20),
                                   hjust = 0),
        axis.text.x = element_text(size = 9,
                                   color = "black",
                                   family = "Montserrat"),
        legend.position = c(0.95, 0.05),
        legend.justification = c("right", "bottom"),
        legend.box.just = "top",
        legend.margin = margin(6, 6, 6, 6),
        legend.direction = "horizontal",
        legend.title = element_text(size = 9,
                                   color = "black",
                                   family = "Montserrat"),
        legend.text = element_text(size = 9,
                                   color = "black",
                                   family = "Montserrat"))

# Importance for major bleeding
varImpBleeding <- ggplot(dfVarImp, 
       aes(y = reorder(label, meanBleeding))) +
  geom_segment(aes(y = reorder(label, meanBleeding), 
                   yend = reorder(label, meanBleeding),
                   x = scaledSvmBleeding,
                   xend = scaledXgbBleeding),
               color = "lightgrey") +
  geom_point(aes(x = scaledSvmBleeding,
                 color = "SVM"),
             alpha = 0.7) +
  geom_point(aes(x = scaledXgbBleeding,
                 color = "XGB"),
             alpha = 0.7) +
  scale_y_discrete(position = "right") +
  scale_x_continuous(position = "top") +
  labs(x = "Estimativa Dimensionada de Desempenho (%)\n") +
  scale_color_manual(name = "Modelo preditor de\nsangramento maior",
                     breaks = c("SVM", "XGB"),
                     values = c("SVM" = "#009fc3", "XGB" = "#6cb43f")) +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 11,
                                   color = "black",
                                   family = "Montserrat",
                                   face = "bold"),
        axis.text.y = element_text(size = 10,
                                   color = "black",
                                   family = "Montserrat",
                                   margin = margin(r = 20),
                                   hjust = 0),
        axis.text.x = element_text(size = 9,
                                   color = "black",
                                   family = "Montserrat"),
        legend.position = c(0.95, 0.05),
        legend.justification = c("right", "bottom"),
        legend.box.just = "top",
        legend.margin = margin(6, 6, 6, 6),
        legend.direction = "horizontal",
        legend.title = element_text(size = 9,
                                   color = "black",
                                   family = "Montserrat"),
        legend.text = element_text(size = 9,
                                   color = "black",
                                   family = "Montserrat"))
```

ROC curves (fig).

```{r, warning=FALSE}

# Packages
library(pROC)
library(mice)
library(tidyverse)

# Data
load("imputation/test.rda")
load("models/predictionLrmMace.rda")
load("models/predictionSvmMace.rda")
load("models/predictionXgbMace.rda")
load("models/predictionLrmBleeding.rda")
load("models/predictionSvmBleeding.rda")
load("models/predictionXgbBleeding.rda")

predictions <- data.frame(mace = complete(impTest, 1)[["mace"]],
                              lrmMace = predictionLrmMace,
                              svmMace = predictionSvmMace[[2]],
                              xgbMace = predictionXgbMace[[2]],
                          bleeding = complete(impTest, 1)[["mace"]],
                              lrmBleeding = predictionLrmBleeding,
                              svmBleeding = predictionSvmBleeding[[2]],
                              xgbBleeding = predictionXgbBleeding[[2]])

# ROC objects
rocLrmMace <- roc(predictions$mace, predictions$lrmMace)
rocSvmMace <- roc(predictions$mace, predictions$svmMace)
rocXgbMace <- roc(predictions$mace, predictions$xgbMace)

rocLrmBleeding <- roc(predictions$bleeding, predictions$lrmBleeding)
rocSvmBleeding <- roc(predictions$bleeding, predictions$svmBleeding)
rocXgbBleeding <- roc(predictions$bleeding, predictions$xgbBleeding)

# ROC curves
rocMace <- ggroc(data = list(rocLrmMace,
                  rocSvmMace,
                  rocXgbMace),
      legacy.axes = TRUE) +
  geom_abline(linetype = "dashed", 
              alpha=0.8,
              color = "grey") +
  scale_color_manual(name = "",
                     labels = c("Regressão", "SVM", "XGB"),
                     values = c("#009fc3", "#6cb43f", "#850a70")) +
  scale_y_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0,0)) +
  scale_x_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0.05,0)) +
  labs(x = "\n1 - Especificidade (%)\n",
       y = "\nSensibilidade (%)\n",
       subtitle = "\nEventos Cardiovasculares\n") +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.6, 0.4),
        legend.justification = c("left", "top"),
        legend.box.just = "top",
        legend.margin = margin(6, 6, 6, 6),
        legend.text = element_text(size = 11,
                                   color = "black",
                                   family = "Montserrat"),
        plot.subtitle = element_text(size = 12,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

rocBleeding <- ggroc(data = list(rocLrmBleeding,
                  rocSvmBleeding,
                  rocXgbBleeding),
      legacy.axes = TRUE) +
  geom_abline(linetype = "dashed", 
              alpha=0.8,
              color = "grey") +
  scale_color_manual(name = "",
                     labels = c("Regressão", "SVM", "XGB"),
                     values = c("#009fc3", "#6cb43f", "#850a70")) +
  scale_y_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0,0)) +
  scale_x_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0.05,0)) +
  labs(x = "\n1 - Especificidade (%)\n",
       y = "\n\nSensibilidade (%)\n",
       subtitle = "\nSangramento Grave\n") +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.6, 0.4),
        legend.justification = c("left", "top"),
        legend.box.just = "top",
        legend.margin = margin(6, 6, 6, 6),
        legend.text = element_text(size = 11,
                                   color = "black",
                                   family = "Montserrat"),
        plot.subtitle = element_text(size = 12,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

#Combining ROC curves
library(gridExtra)
grid.arrange(rocMace, rocBleeding, nrow = 1)

```

Assess and compare ROC curves (test).

```{r, warning=FALSE}

# Packages
library(pROC)
library(mice)
library(tidyverse)

# Data
load("imputation/test.rda")
load("models/predictionLrmMace.rda")
load("models/predictionSvmMace.rda")
load("models/predictionXgbMace.rda")
load("models/predictionLrmBleeding.rda")
load("models/predictionSvmBleeding.rda")
load("models/predictionXgbBleeding.rda")

predictions <- data.frame(mace = complete(impTest, 1)[["mace"]],
                              lrmMace = predictionLrmMace,
                              svmMace = predictionSvmMace[[2]],
                              xgbMace = predictionXgbMace[[2]],
                          bleeding = complete(impTest, 1)[["mace"]],
                              lrmBleeding = predictionLrmBleeding,
                              svmBleeding = predictionSvmBleeding[[2]],
                              xgbBleeding = predictionXgbBleeding[[2]])

# ROC objects
rocLrmMace <- roc(predictions$mace, predictions$lrmMace)
rocSvmMace <- roc(predictions$mace, predictions$svmMace)
rocXgbMace <- roc(predictions$mace, predictions$xgbMace)

rocLrmBleeding <- roc(predictions$bleeding, predictions$lrmBleeding)
rocSvmBleeding <- roc(predictions$bleeding, predictions$svmBleeding)
rocXgbBleeding <- roc(predictions$bleeding, predictions$xgbBleeding)

# Assess c-statistics
rocLrmMace; ci.auc(rocLrmMace)
rocSvmMace; ci.auc(rocSvmMace)
rocXgbMace; ci.auc(rocXgbMace)

rocLrmBleeding; ci.auc(rocLrmBleeding)
rocSvmBleeding; ci.auc(rocSvmBleeding)
rocXgbBleeding; ci.auc(rocXgbBleeding)

# DeLong's test
roc.test(rocLrmMace, rocSvmMace)
roc.test(rocLrmMace, rocXgbMace)

roc.test(rocLrmBleeding, rocSvmBleeding)
roc.test(rocLrmBleeding, rocXgbBleeding)

```

Assess calibration of predictions (fig).

```{r, warning=FALSE}

# Package
#library("devtools"); install_github("BavoDC/CalibrationCurves", dependencies = TRUE)
library("CalibrationCurves")
library(mice)
library(ggmap)
library(tidyverse)

# Data
load("imputation/test.rda")
load("models/predictionLrmMace.rda")
load("models/predictionSvmMace.rda")
load("models/predictionXgbMace.rda")
load("models/predictionLrmBleeding.rda")
load("models/predictionSvmBleeding.rda")
load("models/predictionXgbBleeding.rda")

predictions <- data.frame(mace = complete(impTest, 1)[["mace"]],
                              lrmMace = predictionLrmMace,
                              svmMace = predictionSvmMace[[2]],
                              xgbMace = predictionXgbMace[[2]],
                          bleeding = complete(impTest, 1)[["mace"]],
                              lrmBleeding = predictionLrmBleeding,
                              svmBleeding = predictionSvmBleeding[[2]],
                              xgbBleeding = predictionXgbBleeding[[2]])

# Obtain calibration slope and intercept
calEstimates <- 
  data.frame(slopeLrmMace = val.prob.ci.2(y = predictions$mace,
                             p = predictions$lrmMace)[["Calibration"]][["Slope"]],
             interceptLrmMace = val.prob.ci.2(y = predictions$mace,
                                              p = predictions$lrmMace)[["Calibration"]][["Intercept"]],
             slopeSvmMace = val.prob.ci.2(y = predictions$mace,
                                          p = predictions$svmMace)[["Calibration"]][["Slope"]],
             interceptSvmMace = val.prob.ci.2(y = predictions$mace,
                                              p = predictions$svmMace)[["Calibration"]][["Intercept"]],
             slopeXgbMace = val.prob.ci.2(y = predictions$mace,
                             p = predictions$xgbMace)[["Calibration"]][["Slope"]],
             interceptXgbMace = val.prob.ci.2(y = predictions$mace,
                                              p = predictions$xgbMace)[["Calibration"]][["Intercept"]],
             slopeLrmBleeding = val.prob.ci.2(y = predictions$bleeding,
                                              p = predictions$lrmBleeding)[["Calibration"]][["Slope"]],
             interceptLrmBleeding = val.prob.ci.2(y = predictions$bleeding,
                                                  p = predictions$lrmBleeding)[["Calibration"]][["Intercept"]],
             slopeSvmBleeding = val.prob.ci.2(y = predictions$bleeding,
                                              p = predictions$svmBleeding)[["Calibration"]][["Slope"]],
             interceptSvmBleeding = val.prob.ci.2(y = predictions$bleeding,
                                                  p = predictions$svmBleeding)[["Calibration"]][["Intercept"]],
             slopeXgbBleeding = val.prob.ci.2(y = predictions$bleeding,
                                              p = predictions$xgbBleeding)[["Calibration"]][["Slope"]],
             interceptXgbBleeding = val.prob.ci.2(y = predictions$bleeding,
                                              p = predictions$xgbBleeding)[["Calibration"]][["Intercept"]])

calEstimatesLrmMace <- c(paste("Inclinação: ",
                               calEstimates["Point estimate", "slopeLrmMace"] |> round(2),
                               " (",
                               calEstimates["Lower confidence limit.2.5 %", "slopeLrmMace"] |> round(2),
                               "-",
                               calEstimates["Upper confidence limit.97.5 %", "slopeLrmMace"] |> round(2),
                               ")",
                               sep = ""),
                         paste("Intercepto: ",
                               calEstimates["Point estimate", "interceptLrmMace"] |> round(2),
                               " (",
                               calEstimates["Lower confidence limit.2.5 %", "interceptLrmMace"] |> round(2),
                               "-",
                               calEstimates["Upper confidence limit.97.5 %", "interceptLrmMace"] |> round(2),
                               ")",
                               sep = ""))

calEstimatesSvmMace <- c(paste("Inclinação: ",
                               calEstimates["Point estimate", "slopeSvmMace"] |> round(2),
                               " (",
                               calEstimates["Lower confidence limit.2.5 %", "slopeSvmMace"] |> round(2),
                               "-",
                               calEstimates["Upper confidence limit.97.5 %", "slopeSvmMace"] |> round(2),
                               ")",
                               sep = ""),
                         paste("Intercepto: ",
                               calEstimates["Point estimate", "interceptSvmMace"] |> round(2),
                               " (",
                               calEstimates["Lower confidence limit.2.5 %", "interceptSvmMace"] |> round(2),
                               "-",
                               calEstimates["Upper confidence limit.97.5 %", "interceptSvmMace"] |> round(2),
                               ")",
                               sep = ""))

calEstimatesXgbMace <- c(paste("Inclinação: ",
                               calEstimates["Point estimate", "slopeXgbMace"] |> round(2),
                               " (",
                               calEstimates["Lower confidence limit.2.5 %", "slopeXgbMace"] |> round(2),
                               "-",
                               calEstimates["Upper confidence limit.97.5 %", "slopeXgbMace"] |> round(2),
                               ")",
                               sep = ""),
                         paste("Intercepto: ",
                               calEstimates["Point estimate", "interceptXgbMace"] |> round(2),
                               " (",
                               calEstimates["Lower confidence limit.2.5 %", "interceptXgbMace"] |> round(2),
                               "-",
                               calEstimates["Upper confidence limit.97.5 %", "interceptXgbMace"] |> round(2),
                               ")",
                               sep = ""))

calEstimatesLrmBleeding <- c(paste("Inclinação: ",
                               calEstimates["Point estimate", "slopeLrmBleeding"] |> round(2),
                               " (",
                               calEstimates["Lower confidence limit.2.5 %", "slopeLrmBleeding"] |> round(2),
                               "-",
                               calEstimates["Upper confidence limit.97.5 %", "slopeLrmBleeding"] |> round(2),
                               ")",
                               sep = ""),
                         paste("Intercepto: ",
                               calEstimates["Point estimate", "interceptLrmBleeding"] |> round(2),
                               " (",
                               calEstimates["Lower confidence limit.2.5 %", "interceptLrmBleeding"] |> round(2),
                               "-",
                               calEstimates["Upper confidence limit.97.5 %", "interceptLrmBleeding"] |> round(2),
                               ")",
                               sep = ""))

calEstimatesSvmBleeding <- c(paste("Inclinação: ",
                               calEstimates["Point estimate", "slopeSvmBleeding"] |> round(2),
                               " (",
                               calEstimates["Lower confidence limit.2.5 %", "slopeSvmBleeding"] |> round(2),
                               "-",
                               calEstimates["Upper confidence limit.97.5 %", "slopeSvmBleeding"] |> round(2),
                               ")",
                               sep = ""),
                         paste("Intercepto: ",
                               calEstimates["Point estimate", "interceptSvmBleeding"] |> round(2),
                               " (",
                               calEstimates["Lower confidence limit.2.5 %", "interceptSvmBleeding"] |> round(2),
                               "-",
                               calEstimates["Upper confidence limit.97.5 %", "interceptSvmBleeding"] |> round(2),
                               ")",
                               sep = ""))

calEstimatesXgbBleeding <- c(paste("Inclinação: ",
                               calEstimates["Point estimate", "slopeXgbBleeding"] |> round(2),
                               " (",
                               calEstimates["Lower confidence limit.2.5 %", "slopeXgbBleeding"] |> round(2),
                               "-",
                               calEstimates["Upper confidence limit.97.5 %", "slopeXgbBleeding"] |> round(2),
                               ")",
                               sep = ""),
                         paste("Intercepto: ",
                               calEstimates["Point estimate", "interceptXgbBleeding"] |> round(2),
                               " (",
                               calEstimates["Lower confidence limit.2.5 %", "interceptXgbBleeding"] |> round(2),
                               "-",
                               calEstimates["Upper confidence limit.97.5 %", "interceptXgbBleeding"] |> round(2),
                               ")",
                               sep = ""))

# Calibration curve: lrm for mace
calLrmMace <- ggplot(predictions,
       aes(y = mace,
           x = lrmMace)) +
  geom_abline(linetype = "dashed",
              alpha = 0.5,
              size = 0.5) +
  geom_smooth(method = "lm",
              aes(color = "Calibração logística"),
              se = FALSE) +
  geom_smooth(method = "loess",
              aes(color = "Calibração flexível (Loess)"),
              fill = "lightgray") +
  geom_point(alpha = 0.2) +
  scale_color_manual(name = "",
                     breaks = c("Calibração logística",
                                "Calibração flexível (Loess)"), 
                     values = c("#850A70", "#BE0EA0")) +
  guides(color=guide_legend(override.aes=list(fill=NA))) +
  scale_y_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0,0),
                     limits = c(-0.15, 1.3),
                     breaks = seq(0, 1, by = 0.2)) +
  scale_x_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0.01,0)) +
  labs(x = "\nRisco Predito (%)\n",
       y = "\nProbabilidade Observada (%)\n",
       subtitle = "\nRegressão para Eventos Cardiovasculares\n") +
  annotate("text",
           x = 0.1,
           y = 0.7,
           label = paste(calEstimatesLrmMace[[1]],
                         calEstimatesLrmMace[[2]],
                         sep = "\n"),
           size = 3,
           family = "Montserrat") +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.5, 0.95),
        legend.direction = "horizontal",
        legend.box.just = "center",
        legend.margin = margin(2, 2, 2, 2),
        legend.text = element_text(size = 10,
                                   color = "black",
                                   family = "Montserrat"),
        plot.subtitle = element_text(size = 12,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

# Calibration curve: svm for mace
calSvmMace <- ggplot(predictions,
       aes(y = mace,
           x = svmMace)) +
  geom_abline(linetype = "dashed",
              alpha = 0.5,
              size = 0.5) +
  geom_smooth(method = "lm",
              aes(color = "Calibração logística"),
              se = FALSE) +
  geom_smooth(method = "loess",
              aes(color = "Calibração flexível (Loess)"),
              fill = "lightgray") +
  geom_point(alpha = 0.2) +
  scale_color_manual(name = "",
                     breaks = c("Calibração logística",
                                "Calibração flexível (Loess)"), 
                     values = c("#850A70", "#BE0EA0")) +
  guides(color=guide_legend(override.aes=list(fill=NA))) +
  scale_y_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0,0),
                     limits = c(-0.3, 1.3),
                     breaks = seq(0, 1, by = 0.2)) +
  scale_x_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0.01,0)) +
  labs(x = "\nRisco Predito (%)\n",
       y = "\nProbabilidade Observada (%)\n",
       subtitle = "\nSVM para Eventos Cardiovasculares\n") +
  annotate("text",
           x = 0.09,
           y = 0.7,
           label = paste(calEstimatesSvmMace[[1]],
                         calEstimatesSvmMace[[2]],
                         sep = "\n"),
           size = 3,
           family = "Montserrat") +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.5, 0.95),
        legend.direction = "horizontal",
        legend.box.just = "center",
        legend.margin = margin(2, 2, 2, 2),
        legend.text = element_text(size = 10,
                                   color = "black",
                                   family = "Montserrat"),
        plot.subtitle = element_text(size = 12,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

# Calibration curve: xgb for mace
calXgbMace <- ggplot(predictions,
       aes(y = mace,
           x = xgbMace)) +
  geom_abline(linetype = "dashed",
              alpha = 0.5,
              size = 0.5) +
  geom_smooth(method = "lm",
              aes(color = "Calibração logística"),
              se = FALSE) +
  geom_smooth(method = "loess",
              aes(color = "Calibração flexível (Loess)"),
              fill = "lightgray") +
  geom_point(alpha = 0.2) +
  scale_color_manual(name = "",
                     breaks = c("Calibração logística",
                                "Calibração flexível (Loess)"), 
                     values = c("#850A70", "#BE0EA0")) +
  guides(color=guide_legend(override.aes=list(fill=NA))) +
  scale_y_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0,0),
                     limits = c(-0.15, 1.3),
                     breaks = seq(0, 1, by = 0.2)) +
  scale_x_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0.01,0)) +
  labs(x = "\nRisco Predito (%)\n",
       y = "\nProbabilidade Observada (%)\n",
       subtitle = "\nXGB para Eventos Cardiovasculares\n") +
  annotate("text",
           x = 0.15,
           y = 0.7,
           label = paste(calEstimatesXgbMace[[1]],
                         calEstimatesXgbMace[[2]],
                         sep = "\n"),
           size = 3,
           family = "Montserrat") +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.5, 0.95),
        legend.direction = "horizontal",
        legend.box.just = "center",
        legend.margin = margin(2, 2, 2, 2),
        legend.text = element_text(size = 10,
                                   color = "black",
                                   family = "Montserrat"),
        plot.subtitle = element_text(size = 12,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

# Calibration curve: lrm for bleeding
calLrmBleeding <- ggplot(predictions,
       aes(y = bleeding,
           x = lrmBleeding)) +
  geom_abline(linetype = "dashed",
              alpha = 0.5,
              size = 0.5) +
  geom_smooth(method = "lm",
              aes(color = "Calibração logística"),
              se = FALSE) +
  geom_smooth(method = "loess",
              aes(color = "Calibração flexível (Loess)"),
              fill = "lightgray") +
  geom_point(alpha = 0.2) +
  scale_color_manual(name = "",
                     breaks = c("Calibração logística",
                                "Calibração flexível (Loess)"), 
                     values = c("#257F93", "#31A9C4")) +
  guides(color=guide_legend(override.aes=list(fill=NA))) +
  scale_y_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0,0),
                     limits = c(-0.15, 1.3),
                     breaks = seq(0, 1, by = 0.2)) +
  scale_x_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0.01,0)) +
  labs(x = "\nRisco Predito (%)\n",
       y = "\nProbabilidade Observada (%)\n",
       subtitle = "\nRegressão para Sangramento Grave\n") +
  annotate("text",
           x = 0.085,
           y = 0.7,
           label = paste(calEstimatesLrmBleeding[[1]],
                         calEstimatesLrmBleeding[[2]],
                         sep = "\n"),
           size = 3,
           family = "Montserrat") +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.5, 0.95),
        legend.direction = "horizontal",
        legend.box.just = "center",
        legend.margin = margin(2, 2, 2, 2),
        legend.text = element_text(size = 10,
                                   color = "black",
                                   family = "Montserrat"),
        plot.subtitle = element_text(size = 12,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

# Calibration curve: svm for bleeding
calSvmBleeding <- ggplot(predictions,
       aes(y = bleeding,
           x = svmBleeding)) +
  geom_abline(linetype = "dashed",
              alpha = 0.5,
              size = 0.5) +
  geom_smooth(method = "lm",
              aes(color = "Calibração logística"),
              se = FALSE) +
  geom_smooth(method = "loess",
              aes(color = "Calibração flexível (Loess)"),
              fill = "lightgray") +
  geom_point(alpha = 0.2) +
  scale_color_manual(name = "",
                     breaks = c("Calibração logística",
                                "Calibração flexível (Loess)"), 
                     values = c("#257F93", "#31A9C4")) +
  guides(color=guide_legend(override.aes=list(fill=NA))) +
  scale_y_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0,0),
                     limits = c(-0.3, 1.3),
                     breaks = seq(0, 1, by = 0.2)) +
  scale_x_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0.01,0)) +
  labs(x = "\nRisco Predito (%)\n",
       y = "\nProbabilidade Observada (%)\n",
       subtitle = "\nSVM para Sangramento Grave\n") +
  annotate("text",
           x = 0.1,
           y = 0.7,
           label = paste(calEstimatesSvmBleeding[[1]],
                         calEstimatesSvmBleeding[[2]],
                         sep = "\n"),
           size = 3,
           family = "Montserrat") +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.5, 0.95),
        legend.direction = "horizontal",
        legend.box.just = "center",
        legend.margin = margin(2, 2, 2, 2),
        legend.text = element_text(size = 10,
                                   color = "black",
                                   family = "Montserrat"),
        plot.subtitle = element_text(size = 12,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

# Calibration curve: xgb for bleeding
calXgbBleeding <- ggplot(predictions,
       aes(y = bleeding,
           x = xgbBleeding)) +
  geom_abline(linetype = "dashed",
              alpha = 0.5,
              size = 0.5) +
  geom_smooth(method = "lm",
              aes(color = "Calibração logística"),
              se = FALSE) +
  geom_smooth(method = "loess",
              aes(color = "Calibração flexível (Loess)"),
              fill = "lightgray") +
  geom_point(alpha = 0.2) +
  scale_color_manual(name = "",
                     breaks = c("Calibração logística",
                                "Calibração flexível (Loess)"), 
                     values = c("#257F93", "#31A9C4")) +
  guides(color=guide_legend(override.aes=list(fill=NA))) +
  scale_y_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0,0),
                     limits = c(-0.7, 1.5),
                     breaks = seq(0, 1, by = 0.2)) +
  scale_x_continuous(labels = scales::label_percent(suffix = ""),
                     expand = expansion(0.01,0)) +
  labs(x = "\nRisco Predito (%)\n",
       y = "\nProbabilidade Observada (%)\n",
       subtitle = "\nXGB para Sangramento Grave\n") +
  annotate("text",
           x = 0.2,
           y = -0.4,
           label = paste(calEstimatesXgbBleeding[[1]],
                         calEstimatesXgbBleeding[[2]],
                         sep = "\n"),
           size = 3,
           family = "Montserrat") +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.5, 0.95),
        legend.direction = "horizontal",
        legend.box.just = "center",
        legend.margin = margin(2, 2, 2, 2),
        legend.text = element_text(size = 10,
                                   color = "black",
                                   family = "Montserrat"),
        plot.subtitle = element_text(size = 12,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

# Combine curves
library(gridExtra)
grid.arrange(calLrmMace, calSvmMace, calXgbMace, nrow = 3)
grid.arrange(calLrmBleeding, calSvmBleeding, calXgbBleeding, nrow = 3)


```

Performance estimates of predictions (test).

```{r, warning=FALSE}

# Packages
library(pROC)
library(mice)
library(rms)
library(tidyverse)

# Data
load("imputation/test.rda")
load("models/predictionLrmMace.rda")
load("models/predictionSvmMace.rda")
load("models/predictionXgbMace.rda")
load("models/predictionLrmBleeding.rda")
load("models/predictionSvmBleeding.rda")
load("models/predictionXgbBleeding.rda")

predictions <- data.frame(mace = complete(impTest, 1)[["mace"]],
                              lrmMace = predictionLrmMace,
                              svmMace = predictionSvmMace[[2]],
                              xgbMace = predictionXgbMace[[2]],
                          bleeding = complete(impTest, 1)[["mace"]],
                              lrmBleeding = predictionLrmBleeding,
                              svmBleeding = predictionSvmBleeding[[2]],
                              xgbBleeding = predictionXgbBleeding[[2]])

# ROC objects
rocLrmMace <- roc(predictions$mace, predictions$lrmMace)
rocSvmMace <- roc(predictions$mace, predictions$svmMace)
rocXgbMace <- roc(predictions$mace, predictions$xgbMace)

rocLrmBleeding <- roc(predictions$bleeding, predictions$lrmBleeding)
rocSvmBleeding <- roc(predictions$bleeding, predictions$svmBleeding)
rocXgbBleeding <- roc(predictions$bleeding, predictions$xgbBleeding)

# Ṕerformance measures
measures <- c("specificity", "sensitivity", "ppv", "npv")

rocLrmMace |> 
  pROC::coords(x = "best", best.method="youden", ret = measures, transpose = TRUE) |> 
  round(2)
rocSvmMace |> 
  pROC::coords(x = "best", best.method="youden", ret = measures, transpose = TRUE) |> 
  round(2)
rocXgbMace |> 
  pROC::coords(x = "best", best.method="youden", ret = measures, transpose = TRUE) |> 
  round(2)
rocLrmBleeding |> 
  pROC::coords(x = "best", best.method="youden", ret = measures, transpose = TRUE) |> 
  round(2)
rocSvmBleeding |> 
  pROC::coords(x = "best", best.method="youden", ret = measures, transpose = TRUE) |> 
  round(2)
rocXgbBleeding |> 
  pROC::coords(x = "best", best.method="youden", ret = measures, transpose = TRUE) |> 
  round(2)

```

Comparison of risk classes per outcome (fig).

```{r, warning=FALSE}

library(mice)
library(tidyverse)
library(rms)

# Data from all patients
load("imputation/train.rda")
load("imputation/test.rda")

impAcs <- rbind(complete(impTrain, 1),
          complete(impTest, 1))

# Predict outcomes using lrm
load("models/lrmMace.rda")
load("models/lrmBleeding.rda")

riskMace <- predict(lrmMace,
                    newdata = impAcs,
                    type = "fitted")

riskBleeding <- predict(lrmBleeding,
                        newdata = impAcs,
                        type = "fitted")

# Dataframe with actual outcomes and predicted risk
dfRisk <- data.frame(mace = impAcs$mace,
                     riskMace = riskMace,
                     bleeeding = impAcs$majorBleeding,
                     riskBleeding = riskBleeding,
                     tercileMace = ntile(riskMace, 3),
                     tercileBleeding = ntile(riskBleeding, 3),
                     centileMace = ntile(riskMace, 100),
                     centileBleeding = ntile(riskBleeding, 100))

# Cross classification of risk classes for each outcome
quantile(dfRisk$riskMace, probs = c(1/3, 2/3)) |> round(3)
quantile(dfRisk$riskBleeding, probs = c(1/3, 2/3)) |> round(3)

nTercile <- floor(nrow(impAcs) / 3)

crossRisk <- dfRisk |> 
  group_by(tercileMace,
           tercileBleeding) |> 
  summarise(n = n(),
         percent = n / nTercile * 100,
         percent = ifelse(percent < 5, "", paste(as.character(round(percent)),
                                                 "%",
                                                 sep = "")))

ggplot(crossRisk,
       aes(y = n,
           x = factor(tercileMace,
                      levels = seq(1, 3, by = 1),
                      labels = c(paste("Baixo risco\ncardiovascular\n", "(n = ", nTercile, ")", sep = ""),
                                 paste("Intermediário risco\ncardiovascular\n", "(n = ", nTercile, ")", sep = ""),
                                     paste("Alto risco\ncardiovascular\n", "(n = ", nTercile, ")", sep = ""))),
               fill = factor(tercileBleeding,
                          levels = seq(3, 1, by = -1),
                          labels = c(paste("Alto risco\nhemorrágico\n", "(n = ", nTercile, ")", sep = ""), 
                                     paste("Intermediário risco\nhemorrágico\n", "(n = ", nTercile, ")", sep = ""), 
                                     paste("Baixo risco\nhemorrágico\n", "(n = ", nTercile, ")", sep = ""))),
               label = percent)) +
  geom_col() +
  scale_fill_manual(name = "",
                      labels = c(paste("Alto risco\nhemorrágico\n", "(n = ", nTercile, ")", sep = ""), 
                                     paste("Intermediário risco\nhemorrágico\n", "(n = ", nTercile, ")", sep = ""), 
                                     paste("Baixo risco\nhemorrágico\n", "(n = ", nTercile, ")", sep = "")),
                      values = c("#5BBAD7", "#8CCFE3", "#BDE3EF")) +
  geom_text(size = 4, 
            position = position_stack(vjust = 0.5),
            family = "Montserrat") +
  labs(x = "",
       y = "Número de Pacientes\n") +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = "top",
        legend.direction = "horizontal",
        legend.spacing.x = unit(0.4, "cm"),
        legend.text = element_text(size = 10,
                                   color = "black",
                                   family = "Montserrat"),
        plot.subtitle = element_text(size = 12,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

# Risk differece per centile of mace risk
quantile(dfRisk$riskMace, probs = seq(1, 9, by = 1)/10) |> round(3)

riskTradeoffMace <- dfRisk |> 
  group_by(centileMace,
           tercileBleeding) |> 
  summarise(riskMace = mean(riskMace) * 100,
            riskBleeding = mean(riskBleeding * 100),
            riskDiff = riskMace - riskBleeding) |> 
  ungroup()

ggplot(riskTradeoffMace,
       aes(x = centileMace,
           y = riskDiff,
           color = factor(tercileBleeding,
                          levels = seq(3, 1, by = -1),
                          labels = c("Alto risco hemorrágico", 
                                     "Intermediário risco hemorrágico", 
                                     "Baixo risco hemorrágico")))) +
  geom_hline(alpha = 0.3,
             yintercept = 0,
             linetype = "dashed") +
  geom_smooth(alpha = 0.3,
              fill = "lightgrey") +
  geom_point(alpha = 0.7) +
  scale_color_manual(name = "",
                    labels = c("Alto risco hemorrágico",
                               "Intermediário risco hemorrágico",
                               "Baixo risco hemorrágico"),
                    values = c("#4C0640", "#850A70", "#BE0EA0")) +
  guides(color=guide_legend(override.aes=list(fill=NA))) +
  labs(x = "\nPercentis de Risco Cardiovascular",
       y = "Diferença Absoluta dos Riscos (%)\n") +
  scale_x_continuous(expand = expansion(0.05,0)) +
  scale_y_continuous(expand = expansion(0.05,0)) +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.3, 0.8),
        legend.direction = "vertical",
        legend.box.just = "center",
        legend.margin = margin(2, 2, 2, 2),
        legend.text = element_text(size = 11,
                                   color = "black",
                                   family = "Montserrat"))
  
# Risk difference per centile of bleeding risk
quantile(dfRisk$riskBleeding, probs = seq(1, 9, by = 1)/10) |> round(3)

riskTradeoffBleeding <- dfRisk |> 
  group_by(centileBleeding,
           tercileMace) |> 
  summarise(riskMace = mean(riskMace) * 100,
            riskBleeding = mean(riskBleeding * 100),
            riskDiff = riskMace - riskBleeding) |> 
  ungroup()

ggplot(riskTradeoffBleeding,
       aes(x = centileBleeding,
           y = riskDiff,
           color = factor(tercileMace,
                          levels = seq(3, 1, by = -1),
                          labels = c("Alto risco cardiovascular", 
                                     "Intermediário risco cardiovascular", 
                                     "Baixo risco cardiovascular")))) +
  geom_hline(alpha = 0.3,
             yintercept = 0,
             linetype = "dashed") +
  geom_smooth(alpha = 0.3,
              fill = "lightgrey") +
  geom_point(alpha = 0.7) +
  scale_color_manual(name = "",
                    labels = c("Alto risco cardiovascular",
                               "Intermediário risco cardiovascular",
                               "Baixo risco cardiovascular"),
                    values = c("#185462", "#257F93", "#31A9C4")) +
  guides(color=guide_legend(override.aes=list(fill=NA))) +
  labs(x = "\nPercentis de Risco Hemorrágico",
       y = "Diferença Absoluta entre Riscos (%)\n") +
  scale_x_continuous(expand = expansion(0.05,0)) +
  scale_y_continuous(expand = expansion(0.05,0)) +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.3, 0.20),
        legend.direction = "vertical",
        legend.box.just = "center",
        legend.margin = margin(2, 2, 2, 2),
        legend.text = element_text(size = 11,
                                   color = "black",
                                   family = "Montserrat"))
```

Decision curve analysis (fig, test).

```{r, warning=FALSE}

library(dcurves)
library(rms)
library(mice)
library(tidyverse)

# Outcomes
load("imputation/test.rda")

# Risk predictions
load("models/predictionLrmMace.rda")
load("models/predictionSvmMace.rda")
load("models/predictionXgbMace.rda")
load("models/predictionLrmBleeding.rda")
load("models/predictionSvmBleeding.rda")
load("models/predictionXgbBleeding.rda")

riskPredictions <- data.frame(mace = complete(impTest, 1)[["mace"]],
                              bleeding = complete(impTest, 1)[["majorBleeding"]],
                              predictionLrmMace = predictionLrmMace,
                              predictionSvmMace = predictionSvmMace[[2]],
                              predictionXgbMace = predictionXgbMace[[2]],
                              predictionLrmBleeding = predictionLrmBleeding,
                              predictionSvmBleeding = predictionSvmBleeding[[2]],
                              predictionXgbBleeding = predictionXgbBleeding[[2]])

rm(predictionLrmMace, 
   predictionSvmMace,
   predictionXgbMace,
   predictionLrmBleeding,
   predictionSvmBleeding,
   predictionXgbBleeding)

# DCAs
dcaMace <- dca(formula = mace ~ predictionLrmMace + predictionSvmMace + predictionXgbMace,
    data = riskPredictions) |> 
  plot(smooth = TRUE) +
  labs(x = "\nProbabilidade Limiar (%)\n",
       y = "\nBenefício Líquido (%)\n",
       title = "\nEventos Cardiovasculares\n") +
   scale_color_manual(labels = c("Tratar todos", 
                                 "Tratar ninguém",
                                 "Regressão",
                                 "SVM",
                                 "XGB"), 
                      values = c("#235789",
                                 "#A4B0F5",
                                 "#8FC93A",
                                 "#E6AF2E",
                                 "#E18335")) +
  scale_x_continuous(expand = expansion(0,0.007),
                     limits = c(0,0.25),
                     labels = scales::label_percent(suffix = "")) +
  scale_y_continuous(expand = expansion(0,0),
                     labels = scales::label_percent(suffix = "")) +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.75, 0.7),
        legend.direction = "vertical",
        legend.box.just = "center",
        legend.margin = margin(2, 2, 2, 2),
        legend.text = element_text(size = 11,
                                   color = "black",
                                   family = "Montserrat"),
        plot.title = element_text(size = 12.5,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

dcaBleeding <- dca(formula = bleeding ~ predictionLrmBleeding + predictionSvmBleeding + predictionXgbBleeding,
    data = riskPredictions) |> 
  plot(smooth = TRUE) +
  labs(x = "\nProbabilidade Limiar (%)\n",
       y = "\n\nBenefício Líquido (%)\n",
       title = "\nSangramento Grave\n") +
   scale_color_manual(labels = c("Tratar todos", 
                                 "Tratar ninguém",
                                 "Regressão",
                                 "SVM",
                                 "XGB"), 
                      values = c("#235789",
                                 "#A4B0F5",
                                 "#8FC93A",
                                 "#E6AF2E",
                                 "#E18335")) +
  scale_x_continuous(expand = expansion(0,0.007),
                     limits = c(0,0.25),
                     labels = scales::label_percent(suffix = "")) +
  scale_y_continuous(expand = expansion(0,0),
                     labels = scales::label_percent(suffix = "")) +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.75, 0.7),
        legend.direction = "vertical",
        legend.box.just = "center",
        legend.margin = margin(2, 2, 2, 2),
        legend.text = element_text(size = 11,
                                   color = "black",
                                   family = "Montserrat"),
        plot.title = element_text(size = 12.5,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

library(gridExtra)
grid.arrange(dcaMace, dcaBleeding, nrow = 1)

# Net interventions avoided
interventionsAvoidedMace <- dca(formula = mace ~ predictionLrmMace + predictionSvmMace + predictionXgbMace,
    data = riskPredictions) |> 
  net_intervention_avoided() |> 
  plot(smooth = TRUE) +
  labs(x = "\nProbabilidade Limiar (%)\n",
       y = "\nIntervenções evitadas / 100 pacientes\n",
       title = "\nEventos Cardiovasculares\n") +
   scale_color_manual(labels = c("Regressão",
                                 "SVM",
                                 "XGB"), 
                      values = c("#A4B0F5",
                                 "#8FC93A",
                                 "#E6AF2E")) +
  scale_x_continuous(expand = expansion(0,0),
                     limits = c(0,0.25),
                     labels = scales::label_percent(suffix = "")) +
  scale_y_continuous(expand = expansion(0,0)) +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.7, 0.35),
        legend.direction = "vertical",
        legend.box.just = "center",
        legend.margin = margin(2, 2, 2, 2),
        legend.text = element_text(size = 11,
                                   color = "black",
                                   family = "Montserrat"),
        plot.title = element_text(size = 12.5,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

interventionsAvoidedBleeding <- dca(formula = bleeding ~ predictionLrmBleeding + predictionSvmBleeding + predictionXgbBleeding,
    data = riskPredictions) |> 
  net_intervention_avoided() |> 
  plot(smooth = TRUE) +
  labs(x = "\nProbabilidade Limiar (%)\n",
       y = "\n\nIntervenções evitadas / 100 pacientes\n",
       title = "\nSangramento Grave\n") +
   scale_color_manual(labels = c("Regressão",
                                 "SVM",
                                 "XGB"), 
                      values = c("#A4B0F5",
                                 "#8FC93A",
                                 "#E6AF2E")) +
  scale_x_continuous(expand = expansion(0,0),
                     limits = c(0,0.25),
                     labels = scales::label_percent(suffix = "")) +
  scale_y_continuous(expand = expansion(0,0)) +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(),
        axis.ticks.length = unit(7.5, "pt"),
        axis.title = element_text(size = 12,
                                  family = "Montserrat"),
        axis.text = element_text(size = 10,
                                 family = "Montserrat",
                                 color = "black"),
        legend.position = c(0.7, 0.35),
        legend.direction = "vertical",
        legend.box.just = "center",
        legend.margin = margin(2, 2, 2, 2),
        legend.text = element_text(size = 11,
                                   color = "black",
                                   family = "Montserrat"),
        plot.title = element_text(size = 12.5,
                                    hjust = 0.5,
                                    color = "black",
                                    family = "Montserrat",
                                    face = "bold"))

library(gridExtra)
grid.arrange(interventionsAvoidedMace, interventionsAvoidedBleeding, nrow = 1)

# Clinical consequences for MACE
test_consequences(formula = mace ~ predictionLrmMace + predictionSvmMace + predictionXgbMace,
    data = riskPredictions) |> 
  t()

# Clinical consequences for major bleeding
test_consequences(formula = bleeding ~ predictionLrmBleeding + predictionSvmBleeding + predictionXgbBleeding,
    data = riskPredictions) |> 
  t()
```
